# Pooling the Polls Over an Election Campaign {#campaign}


This example comes from @Jackman2005a.

```{r campaign_setup,message=FALSE}
library("rstan")
library("tidyverse")
library("lubridate")
library("stringr")
library("pscl")
```

House of Representatives - "first preferences""

- https://en.wikipedia.org/wiki/Results_of_the_Australian_federal_election,_2007_(House_of_Representatives)
- https://en.wikipedia.org/wiki/Full_national_lower_house_results_for_the_2004_Australian_federal_election

```{r AustralianElectionPolling}
data("AustralianElectionPolling", package = "pscl")
elections <-
  list(
    `2007` = list(
      date = ymd(20071124),
      ALP = 43.4,
      Lib = 36.3,
      Nat = 5.5,
      Green = 7.8,
      FamilyFirst = 2,
      Dems = 0.7,
      OneNation = 0.3,
      sampleSize = 12419863
    ),
    `2004` = list(
      date = ymd(20041009),
      ALP = 37.6,
      Lib = 40.5,
      Nat = 5.9,
      Green = 7.2,
      FamilyFirst = 2.0,
      Dems = 1.2,
      OneNation = 1.2,
      sampleSize = 11715132      
    )
  )

START_DATE <- elections[["2004"]][["date"]]
  
AustralianElectionPolling <-
  AustralianElectionPolling %>%
  mutate(midDate = as.Date(startDate + difftime(endDate, startDate)),
         ALP_se = ALP * (100 - ALP) / sampleSize,
         time = as.integer(difftime(midDate, START_DATE, units = "days")) + 1L,
         pollster = as.integer(factor(org)))
  
```

For House effects assume mean zero and a standard deviation of 7.5. The standard deviation corresponds
to house effects with a 95% confidence interval of between -15 and 15 (which would be large).

It is expected that most polling movements are $\pm 2$ percentage points. This implies an average scale of

```{r campaign_data}
campaign_data <- within(list(), {
  y <- AustralianElectionPolling$ALP
  s <- AustralianElectionPolling$ALP_se
  time <- AustralianElectionPolling$time
  house <- AustralianElectionPolling$pollster
  H <- max(AustralianElectionPolling$pollster)
  N <- length(y)
  T <- as.integer(difftime(elections[["2007"]][["date"]], elections[["2004"]][["date"]], units = "days")) + 1
  xi_init <- elections[["2004"]][["ALP"]]
  xi_final <- elections[["2007"]][["ALP"]]
  delta_loc <- 0
  tau_scale <- sd(y)
  zeta_scale <- 5
})
```

```{r campaign_mod,results='hide',cache.extra=tools::md5sum("stan/campaign.stan")}
campaign_mod <- stan_model("stan/campaign.stan")
```
```{r echo=FALSE,results='asis'}
campaign_mod
```


```{r campaign_fit,results='hide'}
campaign_fit <- sampling(campaign_mod, data = campaign_data,
                         chains = 1,
                         init = list(list(xi = rep(mean(campaign_data$y), campaign_data$N))))
```
```{r}
campaign_fit
```


```{r campaign_plot_xi}
xi <- summary(campaign_fit, par = "xi")$summary %>%
  as.data.frame() %>%
  rownames_to_column("parameter") %>%
  filter(!str_detect(parameter, ",2\\]")) %>%
  mutate(date = START_DATE + row_number() - 1)

ggplot() +
  geom_ribbon(data = xi, mapping = aes(x = date, ymin = mean - 2 * sd, ymax = mean + 2 * sd), alpha = 0.3) +
  geom_line(data = xi, mapping = aes(x = date, y = mean)) +
  geom_point(data = AustralianElectionPolling,
             mapping = aes(x = midDate, y = ALP, colour = org))

```

