# "Corporatism: hierarchical or "multi-level" model for economic growth in 16 OECD countries" {#corporatism}

```{r corporatism_startup,message=FALSE}
library("rstan")
library("tidyverse")
```

The following program implements a regression model of economic growth among 16 OECD countries, 1971-1984 [@Western1998a, @AlvarezGarrettLange1991a].  The model is hierarchical in that it specifies country-specific coefficients for the following predictors: lagged growth, demand, import price movements, export price movements, leftist government and an intercept.  The magnitudes of the country-specific coefficients are conditional on (time-invariant) extent of labor organization within each country; these regression relationships constitute the 2nd level of the model.

The data come from N=16 countries, and T=14 years (1971:1984) with K=6 covariates at the lowest ("micro") level of the hierarchy, and J=2 covariates (an intercept and the labor organization variable) at the 2nd level.


```{r corporatism}
load("data/corporatism.rda")
corporatism_country <- corporatism %>%
  select(country, labor.org) %>%
  distinct()
  
```

```{r corporatism_mod,results='hide'}
corporatism_mod <- stan_model("stan/corporatism.stan")
```
```{r echo=FALSE,cache=FALSE,results='asis'}
corporatism_mod
```

