# Florida: Learning About an Unknown Proportion from Survey Data {#florida}

```{r florida_setup,message=FALSE}
library("rstan")
```

This is an learning about an unknown proportion from survey data; using survey data to update beliefs about support for Bush in Florida in the 2000 presidential election campaign [@Jackman2004a].

```{r florida_mod,results='hide',cache.extra=tools::md5sum("stan/florida.stan")}
florida_mod <- stan_model("stan/florida.stan")
```
```{r echo=FALSE,cache=FALSE,results='asis'}
florida_mod
```

The prior polls had a mean of 49.1% in support for Bush, with a standard deviation of 2.2%.
The new poll shows 55% support for Bush, with a standard deviation of 2.2%.
```{r florida_data}
florida_data <- list(
  mu_mean = 49.1,
  mu_sd = 2.2,
  y_sd = 2.2,
  y = 55  
)
```

```{r florida_fit,results='hide'}
florida_fit <- sampling(florida_mod, data = florida_data)
```
```{r}
florida_fit
```
```{r include=FALSE}
post_mean <- round(summary(florida_fit)$summary["mu", "mean"], 1)
post_2.5 <- round(summary(florida_fit)$summary["mu", "2.5%"], 1)
post_97.5 <- round(summary(florida_fit)$summary["mu", "97.5%"], 1)
```
After observing the new poll, the mean for the posterior is `r post_mean`, with a 95% credible interval of `r post_2.5`--`r post_97.5`.
