# Legislators: estimating legislators' ideal points from voting histories (roll call data) {#legislators}


```{r legislators_setup,message=FALSE}
library("pscl")
library("tidyverse")
library("forcats")
library("stringr")
library("rstan")
```

Recorded votes in legislative settings (roll calls) are often used to recover the underlying preferences of legislators.  Political scientists analyze roll call data using the Euclidean spatial voting model: each legislator (i = 1, ..., n) has a preferred policy position (xi, a point in low-dimensional Euclidean space), and each vote (j = 1, ..., m) amounts to a choice between "Aye" and a "Nay" locations, qj  and Yj,  respectively. Legislators are assumed to choose on the basis of utility maximization, with utilities (in one-dimension)

In these models, the only observed data are votes, and the analyst wants to 
model those votes as a function of legislator- ($\theta_i$), and vote-specific ($\alpha_i$, $\lambda_i$) parameters.
The vote of legislator $i$ on roll-call $j$ ($y_{i,j}$) is a function of a
the legislator's ideal point ($\theta_i$), the vote's cutpoint ($\alpha_j$), 
and the vote's discrimination ($\lambda_j$):
$$
\begin{aligned}[t]
y_{i,j} &\sim \mathsf{Bernoulli}(\pi_i) \\
\pi_i &= \frac{1}{1 + \exp(-\mu_{i,j})} \\
\mu_{i,j} &= \beta_j \theta_i - \alpha_j
\end{aligned}
$$

Ideal points (like many latent space models) are unidentified. In particular, 
there are three types of invariance:

1. Additive Aliasing
2. Multiplicative Ailising
3. Rotation (Reflection) Invariance


This example models the ideology of the 109th U.S. Senate. 
This data is conveniently included in the `pscl` package:
```{r}
data("s109", package = "pscl")
```
This is not a data frame, see the documentation for its structure.

In our ideal point model, we will use binary responses, but the `s109` data
includes multiple [codes](http://voteview.com/senate109.htm) for respones to 
roll-calls.

---- ----------------------------------------------------------
0    not a member
1    Yea
2    Paired Yea
3    Announced Yea,
4    Announced Nay
5    Paired Nay
6    Nay
7    Present (some Congresses, also not used some Congresses)
8    Present (some Congresses, also not used some Congresses)
6    Nay
9    Not Voting
---- ----------------------------------------------------------

To simplify the analysis, the outcomes will be aggregated to "Yes", "No", and missing values (which 
```{r}
s109_vote_data <- as.data.frame(s109$vote.data) %>%
  mutate(rollcall = paste(session, number, sep = "-"),
         passed = result %in% c("Confirmed", "Agreed To", "Passed"),
         pct_yea = yeatotal / (yeatotal + naytotal),
         unanimous = pct_yea %in% c(0, 1),
         lopsided = pct_yea < 0.35 | pct_yea > 0.65) %>%
  filter(!unanimous) %>%
  mutate(.rollcall_id = row_number())

s109_legis_data <- as.data.frame(s109$legis.data) %>%
  rownames_to_column("legislator") %>%
  mutate(.legis_id = row_number())

s109_votes <- s109$votes %>%
  as.data.frame() %>%
  rownames_to_column("legislator") %>%
  gather(rollcall, vote, -legislator) %>%
  # recode to Yea (TRUE), Nay (FALSE), or missing
  mutate(yea = NA,
         yea = if_else(vote %in% c(1, 2, 3), TRUE, yea),
         yea = if_else(vote %in% c(4, 5, 6), FALSE, yea)) %>%
  filter(!is.na(yea)) %>%
  inner_join(dplyr::select(s109_vote_data, rollcall, .rollcall_id), by = "rollcall") %>%
  inner_join(dplyr::select(s109_legis_data, legislator, party, .legis_id), by = "legislator")

```

Identify votes. Restrict all party line votes to have positive betas:
```{r}
lambda <-
  s109_votes %>%
  group_by(.rollcall_id, party) %>%
  summarise(yea = mean(yea)) %>%
  spread(party, yea) %>%
  ungroup() %>%
  mutate(partyline = "",
         partyline = if_else(R < 0.1 & D > 0.9, "D", partyline),
         partyline = if_else(R > 0.9 & D < 0.1, "R", partyline),
         alpha = if_else(partyline %in% "R", 20,
                          if_else(partyline %in% "D", -20, 0)),
         init = if_else(partyline %in% "R", 1,
                        if_else(partyline %in% "D", -1, 0)))

```

```{r message=FALSE}
mod_ideal_point <- stan_model("ideal_point-4.stan")
```
```{r results='asis'}
mod_ideal_point
```

## No Identification

```{r}
theta <-
  s109_legis_data %>%
  mutate(init = if_else(party == "R", 1,
                        if_else(party == "D", -1, 0)))
```

```{r}
data_ideal_point <-
  within(list(), {
    y <- as.integer(s109_votes$yea)
    y_idx_leg <- as.integer(s109_votes$.legis_id)
    y_idx_vote <- as.integer(s109_votes$.rollcall_id)
    Y_obs <- length(y)
    N <- max(s109_votes$.legis_id)
    K <- max(s109_votes$.rollcall_id)
    # priors
    alpha_loc <- 0
    alpha_scale <- 5.0
    lambda_loc <- rep(0, K)
    lambda_scale <- rep(2.5, K)
    lambda_alpha <- lambda$alpha
  })
```
The prior scale for all the non-fixed $\theta_k$ is weakly informative since the scale is fixed by the two ideal points. However, this may still help regularize some extreme legsilators.
Likewise, a stronger prior on $\alpha$ can help regularize the cutpoints of lopsided votes.

```{r}
init <- list(list( 
  lambda = lambda$init,
  theta = theta$init
))
```

```{r results = 'hide'}
mod_sampling <- sampling(mod_ideal_point,
                         data = data_ideal_point,
                         init = init,
                         chains = 1,
                         iter = 500)
```

Extract the ideal point data:
```{r}
ideal_points <- 
  bind_cols(s109_legis_data,
           as_tibble(summary(mod_sampling, par = "theta")$summary)) %>%
  mutate(legislator = fct_reorder(legislator, mean))
```

```{r}
ggplot(ideal_points, aes(x = legislator, y = mean,
                         ymin = `2.5%`, ymax = `97.5%`, colour = party)) +
  geom_pointrange() +
  coord_flip() + 
  scale_color_manual(values = c(D = "blue", Indep = "green", R = "red"))
```

## Understanding Invariance

Scale invariance:
$$
\begin{aligned}[t]
\mu_{i,j} &= \alpha_j + \lambda_j \theta_i \\
         &= \alpha_j + \left(\frac{\lambda_j}{c}\right) \left(\theta_i c \right) \\
         &= \alpha_j + \lambda^*_j \theta^*_i
\end{aligned}
$$

Addition invariance:
$$
\begin{aligned}[t]
\mu_{i,j} &= \alpha_j + \lambda_j \theta_i \\
          &= \alpha_j - \lambda_j c + \lambda_j c + \lambda_j \theta_i \\
         &= (\alpha_j - \lambda_j c) + \lambda_j (\theta_i + c) \\
         &= \alpha_j^* + \lambda_j \theta^*_i
\end{aligned}
$$

Rotation invariance:
$$
\begin{aligned}[t]
\mu_{i,j} &= \alpha_j + \lambda_j \theta_i \\
          &= \alpha_j  + \lambda_j (-1) (-1) \theta_i \\
          &= \alpha_j  + (-\lambda_j) (-\theta_i) \\          
          &= \alpha_j + \lambda_j^* \theta^*_i
\end{aligned}
$$

Example:
```{r}
theta <- c(-1, -0.5, 0.5, 1)
alpha <- c(1, 0, -1)
lambda <- c(-0.5, 0, 0.5)
y <- matrix(c(1, 0, 1, 0, 0, 1, 1, 1, 1, 0, 1, 1), 3, 4)
k <- 1

list(sum(plogis(y - (alpha + lambda %o% theta))),
     sum(plogis(y - (alpha + -lambda %o% -theta))),
     sum(plogis(y - ((alpha - lambda * k) + lambda %o% (theta + k)))),
     sum(plogis(y - ((alpha + (lambda / k) %o% (theta * k))))))
```

Questions: For each of these: Which types of rotation does it solve?

1. Fix one element of $\lambda$. 
2. Fix one element of $\theta$.
3. Fix one element of $\alpha$.
4. Fix two elements of $\alpha$.
5. Fix two elements of $\theta$.
6. Fix two elements of $\lambda$.


## Identification by Fixing Legislator's Ideal Points

Identification of latent state models can be challenging.
Since we \textit{a priori} know, or expect, that the primary ideological dimension is Liberal-Conservative.
$$
\begin{aligned}[t]
\theta[\text{FRIST (R TN)}] & = 1 \\
\theta[\text{REID (D NV)}] & = -1
\end{aligned}
$$
Otherwise, give a weakly informative prior to the ideal points, and item difficulty and discrimination parameters,
$$
\begin{aligned}[t]
\theta_{k} &\sim \mathsf{Normal}(0, 2.5) \\
\alpha_{k} &\sim \mathsf{Normal}(0, 10.) \\
\lambda_{k} &\sim \mathsf{Normal}(0, 2.5.) 
\end{aligned}
$$

```{r message=FALSE}
mod_ideal_point_1 <- stan_model("ideal_point-2.stan")
```
```{r results='asis'}
mod_ideal_point_1
```

```{r}
theta <-
  s109_legis_data %>%
  mutate(init = if_else(party == "R", 1,
                        if_else(party == "D", -1, 0)),
         # set values
         theta = if_else(legislator == "FRIST (R TN)", 1,
                         if_else(legislator == "REID (D NV)", -1, NA_real_)))
                         
```

```{r}
data_ideal_point_1 <-
  within(list(), {
    y <- as.integer(s109_votes$yea)
    y_idx_leg <- as.integer(s109_votes$.legis_id)
    y_idx_vote <- as.integer(s109_votes$.rollcall_id)
    Y_obs <- length(y)
    N <- max(s109_votes$.legis_id)
    K <- max(s109_votes$.rollcall_id)
    # priors
    alpha_loc <- rep(0, K)
    alpha_scale <- rep(5, K)
    lambda_loc <- rep(0, K)
    lambda_scale <- rep(2.5, K)
    lambda_skew <- rep(0, K)
    N_obs <- sum(!is.na(theta$theta))
    theta_obs_idx <- which(!is.na(theta$theta))
    theta_obs <- theta$theta[!is.na(theta$theta)]
    N_param <- sum(is.na(theta$theta))
    theta_param_idx <- which(is.na(theta$theta))
    theta_loc <- rep(0, N_param)
    theta_scale <- rep(5, N_param)
  })
```

```{r}
init_1 <- list(
  list(theta_param = theta$init[is.na(theta$theta)])
)
```

```{r results = 'hide'}
fit_ideal_1 <- sampling(mod_ideal_point_1,
                        data = data_ideal_point_1,
                        init = init_1,
                        chains = 1,
                        iter = 500)
```

```{r}
ideal_points <- 
  bind_cols(s109_legis_data,
           as_tibble(summary(fit_ideal_1, par = "theta")$summary)) %>%
  mutate(legislator = fct_reorder(legislator, mean))
```

```{r}
ggplot(ideal_points, aes(x = legislator, y = mean,
                         ymin = `2.5%`, ymax = `97.5%`, colour = party)) +
  geom_pointrange() +
  coord_flip() + 
  scale_color_manual(values = c(D = "blue", Indep = "green", R = "red"))
```


### Questions {-}

1. Are there any problems with Rhat or n_eff?
2. Why is Bush's ideal point so uncertain?
3. Why are the ideal points at either end more uncertain?
3. What is the reason for the skew normal distributions? 
4. See Jackman and replicate:

    8. Which Senator is the most liberal, and what is the probability that they are the most liberal?
    9. Which Senator is the median (50%), median of the majority party?
    10. Which bill discriminates the least? 
