# Multivariate Missing Data {#multivarmissing}

```{r multivarmissing,message=FALSE}
library("tidyverse")
library("rstan")
```

This example shows how to impute missing data. See @Stan2016a, Chapter 10 "Missing Data & Partially Known Parameters" for more discussion.

Consider a data set of 10 observations on 3 variables
Only one of the variables, $z$, is completely observed.
The other two variables, x$ and $y$, have a non-overlapping pattern of missing data.

```{r multivarmissing}
multivarmissing <- tribble(
  ~x, ~y, ~z,
1,	NA,	NA,
2,	NA,	4,
3,	NA,	3,
4,	NA,	5,
5,	NA,	7,
NA,	7,	9,
NA,	8,	8,
NA,	9,	11,
NA,	8,	10,
NA,	9,	8)
```

The missing elements of $x$ and $y$ are parameters, and the observed elements of $x$, $y$, and $z$ are data.
These are combined in the `transformed parameters` block, and modeled.

## Separate Regressions

We use $z$ to predict $x$,
and $z$ and $x$ (both observed and imputed) to impute $y$.
$$
\begin{aligned}[t]
x_i &\sim \mathsf{Normal}(\mu_{x,i}, \sigma_x) \\
\mu_{x,i} &= \gamma_1 + \gamma_2 z_i \\
y_i &\sim \mathsf{Normal}(\mu_{y,i}, \sigma_y) \\
\mu_{y,i} &= \beta_1 + \beta_2 y_i + \beta_3 z_i \\
z_i &\sim \mathsf{Normal}(\mu_z, \sigma_z)
\end{aligned}
$$

The parameters are given weakly informative parameters:
$$
\begin{aligned}[t]
\sigma_x,\sigma_y,\sigma_z &\sim \mathsf{HalfCauchy}(0, 5) \\
\gamma_1, \beta_1 &\sim \mathsf{Normal}(0, 10) \\
\gamma_2, \beta_2, \beta_3 &\sim \mathsf{Normal}(0, 2.5)
\end{aligned}
$$
Note that this assumes that $x$, $y$, and $z$ are standardized to have zero mean and unit variance. 

```{r data_multivarmissing}
data_multivarmissing <- within(list(), {
  N <- nrow(multivarmissing)
  x_obs <- multivarmissing$x[!is.na(multivarmissing$x)] %>%
    scale() %>% as.numeric()
  x_obs_idx <- array(which(!is.na(multivarmissing$x)))
  N_x_obs <- length(x_obs_idx)  
  x_miss_idx <- array(which(is.na(multivarmissing$x)))
  N_x_miss <- length(x_miss_idx)
  y_obs <- multivarmissing$y[!is.na(multivarmissing$y)] %>%
    scale() %>% as.numeric()    
  y_obs_idx <- array(which(!is.na(multivarmissing$y)))
  N_y_obs <- length(y_obs_idx)  
  y_miss_idx <- array(which(is.na(multivarmissing$y)))
  N_y_miss <- length(y_miss_idx)
  z_obs <- multivarmissing$z[!is.na(multivarmissing$z)] %>%
    scale() %>% as.numeric()
  z_obs_idx <- array(which(!is.na(multivarmissing$z)))
  N_z_obs <- length(z_obs_idx)
  z_miss_idx <- array(which(is.na(multivarmissing$z)))
  N_z_miss <- length(z_miss_idx)
  alpha_loc <- 0
  alpha_scale <- 10
  beta_loc <- rep(0, 3)
  beta_scale <- c(10, 2.5, 2.5)
  gamma_loc <- rep(0, 2)
  gamma_scale <- c(10, 2.5)
  sigma_x_scale <- 5
  sigma_y_scale <- 5
  sigma_z_scale <- 5
})
```

```{r mod_multivarmissing,cache.extra=tools::md5sum("stan/multivarmissing.stan")}
mod_multivarmissing <- stan_model("stan/multivarmissing2.stan")
```
```{r}
mod_multivarmissing
```

```{r fit_multivarmissing,results='hide'}
fit_multivarmissing <- 
  sampling(mod_multivarmissing, data = data_multivarmissing)
```
```{r}
fit_multivarmissing
```

## Multivariate Normal

```{r data_multivarmissing2}
data_multivarmissing2 <- within(list(), {
  N <- nrow(multivarmissing)
  K <- ncol(multivarmissing)

  mu_loc <- rep(0, 3)
  mu_scale <- rep(0, 10)
  Sigma_scale_scale <- 5
  Sigma_corr_L_eta <- 2
})
```

```{r mod_multivarmissing,cache.extra=tools::md5sum("stan/multivarmissing2.stan")}
mod_multivarmissing2 <- stan_model("stan/multivarmissing2.stan")
```
```{r}
mod_multivarmissing2
```

```{r fit_multivarmissing,results='hide'}
fit_multivarmissing <- 
  sampling(mod_multivarmissing2, data = data_multivarmissing2)
```
```{r}
fit_multivarmissing
```
