# Reagan: linear regression with AR(1) disturbances {#reagan}

```{r reagan_setup,message=FALSE,cache=FALSE}
library("tidyverse")
library("rstan")
```


Ninety-six monthly observations on presidential job approval ratings for Ronald Reagan are modeled via linear regression, with a correction for first-order serial correlation among the disturbances.  Note the marginal model for the first observation, and the conditioning on the lagged observation for months 2 through 96.  A uniform prior over the stationary (-1,1) interval is employed for the residual AR(1) parameter.

$$
\begin{aligned}[t]
y_i &= \mu_i + \epsilon_i + \theta \epsilon_{i - 1}  ,\\
\mu_i &= \alpha + x_i' \beta , \\
\epsilon_i &\sim \mathsf{Normal}(0, \sigma^2) ,
\end{aligned}
$$
for $i \in 1, \dots, N$.
Weakly informative priors for each parameter are used,
$$
\begin{aligned}[t]
\alpha &\sim \mathsf{Normal}(0, 10), \\
\beta_k &\sim \mathsf{Normal}(0, 2.5), & k \in 1, \dots, K, \\
\sigma &\sim \mathsf{HalfCauchy}(0, 5), \\
\theta &= 2 \theta^*  - 1 , \\
\theta^* &\sim \mathsf{Beta}(1, 1)  .
\end{aligned}
$$

```{r ReaganApproval}
(load("data/ReaganApproval.rda"))
ReaganApproval
```


```{r reagan_data}
reagan_data <- within(list(), {
  y <- ReaganApproval$app
  N <- length(y)
  X <- model.matrix(~ 0 + infl + unemp, data = ReaganApproval) %>% scale()
  K <- ncol(X)
  alpha_loc <- 0
  alpha_scale <- 10
  beta_loc <- rep(0, K)
  beta_scale <- rep(2.5 * sd(y), K)
  sigma_scale <- 5 * sd(y)
  theta_a <- 1
  theta_b <- 1
})
```

```{r mod_regar1,cache.extra=tools::md5sum("stan/regar1.stan")}
mod_regar1 <- stan_model("stan/regar1.stan")
```
```{r}
mod_regar1
```

```{r reagan_fit,results='hide'}
reagan_fit <- sampling(mod_regar1, data = reagan_data)
```

```{r reagan_fit_summary}
summary(reagan_fit, par = c("alpha", "beta", "theta", "sigma"))$summary
```


## Cochrane-Orcutt/Prais-Winsten

An AR(1) error model can also be estimated via partial-differencing as 
in Cochrane-Orcutt/Prais-Winsten estimation.

