# Political Sophistication: item-response modeling with mixed data types {#sophistication}

```{r sophistication_setup,message=FALSE,cache=FALSE}
library("tidyverse")
library("rstan")
```


## Data

```{r PoliticalSophistication}
(load("data/PoliticalSophistication.rda"))
```

As part of a survey of French public opinion , 2,148 respondents were administered a series of 19 items assessing their knowledge of political leaders, political institutions, constitutional provisions, and the policies of the political parties [@GrunbergMayerSniderman2002a].
Each response is coded "correct" (1) or "incorrect" (0), and is modeled via a two-parameter item-response model, with a logistic link function; each respondent's level of political sophistication is the latent trait.

In addition, at the end of the thirty minute phone interview, the interviewer assigned a score for each respondent's level of political information (based on their impressions of the respondents formed over the course of the entire interview) on a zero to twenty scale. These responses are modeled via a linear regression, with each respondent's latent trait appearing as an unobserved predictor, and an intercept specific to each interviewer (modeled hierarchically in the code below). To uniquely orient the latent trait (higher values corresponding to more political sophistication),  the interviewer ratings are constrained to positively discriminate with respect to the latent trait (see the constraint on the prior for gamma).


## Model

The survey data consists of 20 items, $y_1, \dots, y_20$. The first 19 items, $y_1, \dots, y_19$ binary responses to political information questions.
The final item, $y_20$, is a political sophistication score (0--20) assigned by the interviewer.

Let $y_{i,j}$ be the response of respondent $i \in 1, \dots, N$ to question $j \in 1, \dots, 20$.
$$
\begin{aligned}[t]
y_{i,j} &\sim \mathsf{Bernoulli}(\mathsf{Logit}^{-1}(\beta_j \xi_i - \alpha_j)) 
\end{aligned}
$$
for $j \in 1, \dots, j$.
Item 20 is modeled as
$$
\begin{aligned}[t]
y_{20,j} &\sim \mathsf{Normal}(\theta_i, \sigma^2) , \\
\theta_i &= \gamma \xi_i + \nu_{m[i]} .
\end{aligned}
$$
Since the question is assigned by the interviewer, $\theta_i$ is a linear function of a the latent score of the respondent ($\xi_i$) and an interviewer specific random effect, $\nu_{m[i]}$, where $m[i]$ means that $i$ was interviewed by interviewer $m \in 1, \dots, M$.
The interviewer effects are given a prior distribution, 
$$
\nu_m \sim \mathsf{Normal}(0, \tau)  .
$$

To fix scale and location invariances, the respondents' abilities are given a standard normal distribution, 
$$
\xi_m \sim \mathsf{Normal}(0, 1) .
$$
Since higher interviewer assessments should be associated with a higher latent political knowledge score, the rotation invariance is resolved by restricting the coefficient for the respondents to be positive, 
$$
\gamma \sim \mathsf{HalfNormal}(0, 2.5 s_z) .
$$

The remaining parameters are assigned weakly informative priors.

$$
\begin{aligned}[t]
\tau &\sim \mathsf{HalfCauchy}(0, 5 s_{y_{20}}) , \\
\sigma &\sim \mathsf{HalfCauchy}(0, 5 s_{y_{20}})  , \\
\delta &\sim \mathsf{Normal}(10, 10 s_{y_{20}}) , \\
\beta_k &\sim \mathsf{Normal}(0, 2.5) , \\
\alpha_k &\sim \mathsf{Normal}(0, 10) .
\end{aligned}
$$
where $s_{y_{20}}$ is the scale for $y_20$.
We could use the empirical standard deviation of $y_20$, or an a-priori measure.
A value of $s_{y_{20}} = 21 / 4$, would place 95% of the mass of a normal distribution between 0 and 20.

```{r mod_sophistication,results='hide'}
mod_sophistication <- stan_model("stan/sophistication.stan")
```
```{r echo=FALSE,results='asis'}
mod_sophistication
```


## Estimation

```{r data_sophistication}
y_scale <- 21 / 4
  
data_sophistication <- within(list(), {
  y_bern <- t(as.matrix(select(PoliticalSophistication, Q1:Q19)))
  N <- ncol(y_bern)
  K <- nrow(y_bern)
  y_norm <- PoliticalSophistication$Q20
  interviewer <- PoliticalSophistication$interviewer
  J <- max(interviewer)
  # priors
  sigma_scale <- 5 * y_scale
  xi_loc <- 0
  xi_scale <- 1
  alpha_loc <- 0
  alpha_scale <- 5
  beta_loc <- 0
  beta_scale <- 2.5
  gamma_scale <- 2.5 * y_scale
  # priors for interviewer effects
  tau_scale <- y_scale
  delta_loc <- 10
  delta_scale <- y_scale
})
```

```{r}
fit_sophistication <- sampling(mod_sophistication, data = data_sophistication, init = 0, chains = 1)
```


## Questions / Extensions

- An alternative parameterization would place the political sophistication on the same 0-20 scale as `Q20`. 
- Model Q20 as an ordinal variable instead of a continuous variable.
