# Truncation: How does Stan deal with truncation?

```{r truncated_setup,message=FALSE}
library("tidyverse")
library("rstan")
```

Suppose we observed $y = (1, \dots, 9)$,[^truncated-source]
```{r}
y <- 1:9
```
These observations are drawn from a population distributed normal with unknown mean ($\mu$) and variance ($\sigma^2$), with the constraint that $y < 10$,
$$
\begin{aligned}[t]
y_i &\sim \mathsf{Normal}(\mu, \sigma^2) I(-\infty, 10) .
\end{aligned}
$$

With the censoring taken into account, the log likelihood is 
$$
\log L(y; \mu, \sigma) = \sum_{i = 1}^n \left( \log \phi(y_i; \mu, \sigma^2) - \log\Phi(y_i; \mu, \sigma^2) \right)
$$
where $\phi$ is the normal distribution PDF, and $\Phi$ is the normal distribution $

The posterior of this model is not well identified by the data, so the mean, $\mu$, and scale, $\sigma$, are given informative priors based on the data,
$$
\begin{aligned}[t]
\mu &\sim \mathsf{Normal}(\bar{y}, s_y) ,\\
\sigma &\sim \mathsf{HalfCauchy}(0, s_y) .
\end{aligned}
$$
where $\bar{y}$ is the mean of $y$, and $s_y$ is the standard deviation of $y$. Alternatively, we could have standardized the data prior to estimation.

## Stan Model

See @Stan2016a, Chapter 11 "Truncated or Censored Data" for more on how Stan handles truncation and censoring. 
In Stan the `T` operator used in sampling statement,
```
y ~ distribution(...) T[upper, lower];
```
is used to adjust the log-posterior contribution for truncation.


```{r truncate_mod,results='hide'}
truncate_mod <- stan_model("stan/truncated.stan")
```
```{r echo=FALSE,results='asis',cache=FALSE}
truncate_mod
```



## Estimation

```{r truncate_data}
truncate_data <- within(list(), {
  y <- y
  N <- length(y)
  U <- 10
  mu_mean <- mean(y)
  mu_scale <- sd(y)
  sigma_scale <- sd(y)
})
```


```{r truncate_fit1,results='hide',message=FALSE}
truncate_fit1 <- sampling(truncate_mod, data = truncate_data)
```
```{r}
truncate_fit1
```

We can compare these results to that of a model in which the truncation is not taken into account:
$$
\begin{aligned}[t]
y_i &\sim \mathsf{Normal}(\mu, \sigma^2), \\
\mu &\sim \mathsf{Normal}(\bar{y}, s_y) ,\\
\sigma &\sim \mathsf{HalfCauchy}(0, s_y) .
\end{aligned}
$$

```{r truncate_mod2,results='hide'}
truncate_mod2 <- stan_model("stan/normal.stan")
```
```{r echo=FALSE,results='asis',cache=FALSE}
truncate_mod2
```

```{r truncate_fit2,results='hide'}
truncate_fit2 <-
  sampling(truncate_mod2, data = truncate_data)
```
```{r}
truncate_fit2
```

We can compare the densities for $\mu$ and $\sigma$ in each of these models.
```{r truncted_plot_density}
plot_density <- function(par) {
  bind_rows(
    tibble(value = rstan::extract(truncate_fit1, par = par)[[1]],
            model = "truncated"),
    tibble(value = rstan::extract(truncate_fit2, par = par)[[1]],
            model = "non-truncated")
  ) %>%
    ggplot(aes(x = value, colour = model, fill = model)) +
    geom_density(alpha = 0.3) +
    labs(x = eval(bquote(expression(.(as.name(par)))))) +
    theme(legend.position = "bottom")
}
```
```{r truncate_plot_density_mu,fig.cap="Posterior density of $\\mu$ when estimated with and without truncation"}
plot_density("mu")
```
```{r truncate_plot_density_sigma,fig.cap="Posterior density of $\\sigma$ when estimated with and without truncation"}
plot_density("sigma")
```


## Questions

1. How are the densities of $\mu$ and $\sigma$ different under the two models? Why are they different?
2. Re-estimate the model with improper uniform priors for $\mu$ and $\sigma$. How do the posterior distributions change?
3. Suppose that the truncation points are unknown. Write a Stan model and estimate. See @Stan2016a, Section 11.2 "Unknown Truncation Points" for how to write such a model. How important is the prior you place on the truncation points? 

[^truncated-source]: This example is derived from Simon Jackman. "Truncation: How does WinBUGS deal with truncation?" *BUGS Examples*, 2007-07-24,
 [URL](https://web-beta.archive.org/web/20070724034035/http://jackman.stanford.edu:80/mcmc/SingleTruncation.odc).
