# Truncation: How does Stan deal with truncation?

See @Stan2016a, Chapter 11 "Truncated or Censored Data" for more on how Stan handles truncation and censoring.
```{r message=FALSE}
library("tidyverse")
library("rstan")
```


Assume we have the observations, $y = 1,...,9$, from a Normal population with unknown mean and variance, subject to the constraint that $y < 10$,
$$
\begin{aligned}[t]
y &\sim \mathsf{Normal}(\mu, \sigma^2) I(-\infty, 10) .
\end{aligned}
$$

Ignoring the constraint, the MLEs for the mean and variance are 5 and 6.67; with the constraint taken into account, each observation makes likelihood contribution
$$
f (y; m, s_2)/F ((k - m)/s),
$$
where $k$ is the truncation point (in this case, 10), and the MLEs of $m, s_2$  are 5.32 and 8.28.

The posterior of this model is not well identified by the data, so the mean, $\mu$, and scale, $\sigma$, are given informative priors based on the data,
$$
\begin{aligned}[t]
\mu &\sim \mathsf{Normal}(\bar{y}, s_y) ,\\
\sigma &\sim \mathsf{HalfCauchy}(0, s_y) .
\end{aligned}
$$
where $\bar{y}$ is the mean of $y$, and $s_y$ is the standard deviation of $y$.

```{r truncation_mod,results='hide'}
truncation_mod <- stan_model("stan/SingleTruncation.stan")
```
```{r echo=FALSE,results='asis',cache=FALSE}
truncation_mod
```

```{r truncation_data}
truncation_data <- within(list(), {
  y <- 1:9
  N <- length(y)
  U <- 10
  mu_mean <- mean(y)
  mu_scale <- sd(y)
  sigma_scale <- sd(y)
})
```



```{r truncation_fit,results='hide',message=FALSE}
truncation_fit <- sampling(truncation_mod, data = truncation_data)
```
```{r}
truncation_fit
```

We can compare these results to that of a model in which the truncation is not taken into account:
$$
\begin{aligned}[t]
y_i &\sim \mathsf{Normal}(\mu, \sigma^2), \\
\mu &\sim \mathsf{Normal}(\bar{y}, s_y) ,\\
\sigma &\sim \mathsf{HalfCauchy}(0, s_y) .
\end{aligned}
$$

```{r truncation_mod2,results='hide'}
truncation_mod2 <- stan_model("stan/normal.stan")
```
```{r echo=FALSE,results='asis',cache=FALSE}
truncation_mod2
```

```{r truncation_fit2,results='hide'}
truncation_fit2 <-
  sampling(truncation_mod2, data = truncation_data)
```
```{r}
truncation_fit2
```


