# Turnout: logit/probit models for binary data {#turnout}

```{r turnout_setup,message=FALSE,cache=FALSE}
library("tidyverse")
library("rstanarm")
library("rstan")
```

## Data

The data comprise the first 2,000 (of 15,000+) observations in the 1992 [American National Election Studies](http://www.electionstudies.org/) (ANES).
These data are included in the **Zelig** package as `turnout` and analyzed in @KingTomzWittenberg2000a.
```{r turnout)}
data("turnout", package = "Zelig")
glimpse(turnout)
```
We will model voting turnout as a function of covariates (age, education, income, race).
```{r turnout_formula}
turnout_formula <- vote ~ poly(age, 2) + educate + income + race
```

## Logit Model

Let $y_i \in \{0, 1\}$ be the decision to vote by respondent $i$ for $i \in 1, \dots, n$,
$$
\begin{aligned}[t]
y_i &\sim \mathsf{Bernoulli}(\pi_i) , \\
\pi_i &= \frac{1}{1 + e^{-\eta_i}} , \\
\eta_i &= \alpha + x_i \beta,
\end{aligned}
$$
where $x_i$ is a vector of covariates.
The regression parameters, $\alpha$ and $\beta$, are given weakly informative priors on the logit scale,
$$
\begin{aligned}[t]
\alpha &\sim \mathsf{Normal}(0, 16) , \\
\beta_k &\sim \mathsf{Normal}(0, 4) .
\end{aligned}
$$

The logit model in Stan is
```{r turnout_mod_logit,results='hide',cache.extra=tools::md5sum("stan/logit.stan")}
turnout_mod_logit <- stan_model("stan/logit.stan")
```
```{r results='asis',echo=FALSE}
turnout_mod_logit
```

```{r turnout_data}
turnout_data <- within(list(), {
  N <- nrow(turnout)
  X <- scale(model.matrix(update(turnout_formula, . ~ 0 + .),
                                 data = turnout))
  K <- ncol(X)
  y <- turnout$vote
  alpha_loc <- 0
  alpha_scale <- 10
  beta_loc <- rep(0, K)
  beta_scale <- rep(2.5, K)
})
```

```{r turnout_fit_logit,results='hide'}
turnout_fit_logit <- sampling(turnout_mod_logit, data = turnout_data)
```
```{r}
turnout_fit_logit
```

## Probit Model

The only difference between the logit and probit models are in the link function.
The probit model uses the normal distribution CDF function instead of the inverse logit.
$$
\begin{aligned}[t]
\pi_i &= \Phi(\eta_i) ,
\end{aligned}
$$
where $\Phi$ is the standard normal distribution CDF.
The priors for the probit model are adjusted by a facto of 1.6, corresponding to `dnorm(0) / dlogis(0)`,
$$
\begin{aligned}[t]
\alpha &\sim \mathsf{Normal}(0, 16), \\
\beta_k &\sim \mathsf{Normal}(0, 4) .
\end{aligned}
$$
```{r}
turnout_data <- within(turnout_data, {
  alpha_scale <- 16
  beta_scale <- rep(4, K)
})
```

The probit model in Stan is
```{r turnout_mod_probit,results='hide',cache.extra=tools::md5sum("stan/probit.stan")}
turnout_mod_probit <- stan_model("stan/probit.stan")
```
```{r results='asis',echo=FALSE}
turnout_mod_probit
```

Fit the model.
```{r turnout_fit_probit,results='hide'}
turnout_fit_probit <-
  sampling(turnout_mod_probit,
           data = turnout_data)
```
```{r}
turnout_fit_probit
```

## rstanarm

These models can also be estimated with the  **rstanarm** function `stan_glm`.
```{r turnout_fit_logit2,results='hide'}
turnout_fit_logit2 <-
  stan_glm(turnout_formula,
           family = binomial(),
           data = turnout)
```
```{r}
turnout_fit_logit2
```

For the probit model, the priors need to be reduced.
```{r turnout_fit_probit2,results='hide'}
turnout_fit_probit2 <-
  stan_glm(turnout_formula,
           family = binomial(link = "probit"),
           prior_intercept = normal(location = 0, scale = 8, autoscale = TRUE),
         prior = normal(location = 0, scale = 1.5, autoscale = TRUE),
         init = 0, chains = 1,
         data = turnout)
```
```{r}
turnout_fit_probit2
```

## Questions {-}

1.  Estimate the posterior distribution of the average marginal effect for each covariate using the methods in @HanmerKalkan2012a.
1.  Estimate the expected percent correctly predicted (ePCP) and percent correctly predicted (PCP) as discussed in the [cosponsorship data](#cosponsor).
1.  Estimate a generalized logit model to allow for asymmetric effects of covariates as described in the [General Beetles](#genbeetles) example and @Nagler1994a. Describe the substantive meaning of this generalization.
1.  Is there any difference in the fit of the probit and logit? Which observations fit better or worse.
1.  Estimate the model with a Student's $t$ CDF instead of logit or probit. This is called a robit model @Liu2005a. Compare its fit and estimates to those of the logit. Which observations fit better or worse? Describe the substantive meaning of this generalization.
