# House of Commons elections: modeling with the multivariate Student-$t$ density {#uk92}


```{r uk92_setup,message=FALSE,cache=FALSE}
library("tidyverse")
library("rstan")
```

The following analysis comes from a data set on United Kingdom House of Commons elections, used in @KatzKing1999a, and re-analyzed in @TomzTuckerWittenberg2002a.  The data for this example consist of vote proportions from 1992 for 521 constituencies, for the Labor party, the Conservative Party, and the Liberal-Alliance, transformed to a 521 by 2 matrix of log-odds ratios (with the Liberal-Alliance in the denominator), as is common in the analysis of multinomial or "compositional" data [@Aitchison1982a].  @KatzKing1999a noted that the distribution of the log-odds ratios appear to be heavy-tailed relative to the normal, and model the data with a multivariate t-distribution with unknown degrees of freedom.

```{r}
data("UKHouseOfCommons", package = "pscl")
```


This specification includes covariates for the conditional means of the two log odds ratios.  Six covariates are used: 

- a constant
- the lagged values (from the previous election) of the two log odds ratios
- an indicator for whether the Conservative candidate is the incumbent
- the incumbency status of the Labor candidate
- the incumbency status of the Liberal/Alliance candidate.  

$$
\begin{aligned}[t]
y_i &\sim \mathsf{StudentT}(\nu, \alpha + x' \beta, \Sigma) & i \in 1, \dots, N,
\end{aligned}
$$
where $\Sigma$ is a correlation matrix.
$y_i$ are log-odds.

For identifiction, as in a logit regression, either the intercept or scale must be fixed. In this case, $\Sigma$ is a correlation matrix.

Weakly informative priors are used for the regression parameters. 
The degrees of freedom of the multivariate Student t distribution is a parameter, and given a weakly informative Gamma distribution that puts most of the prior density between 3 and 40 [@JuarezSteel2010a],
$$
\begin{aligned}[t]
\alpha &\sim  \mathsf{Normal}(0, 10) , \\
\beta_p &\sim \mathsf{Normal}(0, 2.5), & p \in 1, \dots, P , \\
\Sigma &\sim \mathsf{LkjCorr}(\eta) , \\
\nu &\sim \mathsf{Gamma}(2, 0.1) .
\end{aligned}
$$

```{r uk92,cache.extra=tools::md5sum("data/uk92.rda")}
(load("data/uk92.rda"))
```


```{r uk92_data}
uk92_data <- within(dget("data/uk92.R"), {
  y <- as.matrix(dplyr::select(UKHouseOfCommons, y1, y2))
  X <- model.matrix()
  N <- nrow(y)
  K <- ncol(y)
  P <- ncol(X)
  alpha_location <- rep(0, K)
  alpha_scale <- rep(10, K)
  beta_location <- matrix(0, K, P)
  beta_scale <- matrix(2.5, K, P)
  Sigma_shape <- 1
})
```

```{r results='hide',cache.extra=tools::md5sum("stan/uk92.stan")}
uk92_mod <- stan_model("stan/uk92.stan")
```
```{r results='asis',echo=FALSE}
uk92_mod
```

```{r uk92_fit,results='hide'}
uk92_fit <- sampling(uk92_mod, data = uk92_data)
```
```{r}
summary(uk92_fit, par = c("nu", "a", "b", "Sigma"))$summary
```
