# Undervoting for President, by Race: Difference in Two Binomial Proportions {#undervote}

```{r undervote_setup,message=FALSE,cache=FALSE}
library("tidyverse")
library("rstan")
```

Does undervoting for the US president differ by race?
Intentional undervoting is when a voter chooses not to cast vote for
an item on a ballot.

@TomzHouweling2003a analyze this phenomenon using two surveys:

-   Voter News Service (VNS) exit poll for the 1992 election
-   American National Election Studies (ANES) for the 1964--2000 elections

Each of these surveys asked voters whether they voted for president, as well as the race of the respondents.
The results of these surveys is contained in the `undervote` data frame.
The column `undervote` is the number of respondents who reported voting but not voting for president.

```{r undervote}
undervote <- tribble(
  ~survey, ~race, ~n, ~undervote,
  "VNS", "black", 6537, 26,
  "VNS", "white", 44531, 91,
  "ANES", "black", 1101, 10,
  "ANES", "white", 9827, 57
  )
```

```{r echo=FALSE,results='asis'}
undervote %>%
  mutate(`Survey` = survey,
         `Race` = race,
         `No. Voted` = n,
         `Didn't vote for president` = undervote) %>%
  knitr::kable()
```

We are interested in analyzing the difference in proportions for each of these surveys independently.
We will model the proportions of each race and survey,
$$
\begin{aligned}[t]
y_i &\sim \mathsf{Binomial}(n_i, \pi_i) ,
\end{aligned}
$$
where
$$
i \in \{ (\text{VNS},\text{black}), (\text{VNS},\text{white}),  (\text{ANES},\text{black}),  (\text{ANES},\text{white}) \} .
$$

We will model the proportions independently by assigning them identical independent uninformative priors,

$$
\begin{aligned}[t]
\pi_i &\sim \mathsf{Beta}(1, 1) .
\end{aligned}
$$
The racial differences in undervoting in each survey are auxiliary quantities,
$$
\begin{aligned}[t]
\delta_{\text{VNS}} &= \pi_{\text{VNS},\text{black}} - \pi_{\text{VNS},\text{white}} ,\\
\delta_{\text{ANES}} &= \pi_{\text{ANES},\text{black}} - \pi_{\text{ANES},\text{white}} . \\
\end{aligned}
$$
We are also interested in the probability that black undervoting is greater than white undervoting, $\Pr(\delta_j) > 0$, in each survey.

```{r undervote_mod,results='hide',cache.extra=tools::md5sum("stan/undervote.stan")}
undervote_mod <- stan_model("stan/undervote.stan")
```
```{r results='asis',echo=FALSE}
undervote_mod
```

```{r undervote_data}
# this analysis depends on the order of the data frame
undervote_data <-
  list(y = undervote$undervote,
       n = undervote$n,
       N = nrow(undervote),
       pi_a = rep(1, 4),
       pi_b = rep(1, 4))
```

```{r undervote_fit,results='hide'}
undervote_fit <- sampling(undervote_mod, data = undervote_data)
```
```{r}
undervote_fit
```

## References {-}

Simon Jackman, "[Undervoting for President, by Race: difference in two binomial proportions](https://web-beta.archive.org/web/20070724034102/http://jackman.stanford.edu:80/mcmc/undervote.odc)", *BUGS Examples* 2007-07-24.
